Hierarchical Phrase-Based Grammar Extraction in Joshua

### 2009-12-01 papers/mt_marathon10_SuffixArray/paper.tex
Work on implementation section
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1203 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

The remainder of this paper is structured as follows. Section \ref{related-work} briefly overviews the grammar extraction mechanisms employed by other prominent open source machine translation systems. Section \ref{data-structures} reviews the data structures that are used for efficient run-time calculation of SCFG rules. Section \ref{algorithms} presents two core algorithms of this implementation that differ from those presented in \newcite{Lopez2008}. Finally, section \ref{using-joshua} provides details for configuring and running the Joshua grammar extraction tools.

@@

\section{Data Structures}
\label{data-structures}

\subsection{Symbol Table}

\subsection{Suffix Array}

\subsection{Prefix Tree with Suffix Links}

\subsection{Memory Mapping}

\section{Algorithms}
\label{algorithms}

\subsection{Caching}
Discuss caching procedure

\subsection{Query Intersection}
\label{query-intersection}
Discuss query intersect

@@

%\end{algorithm}
\caption{Query intersection algorithm implemented in Joshua. This algorithm is adapted from a corrected version (Lopez, p.c.) of query intersection {\protect \cite{Lopez2008}}.}
\end{figure}

>>>>

The remainder of this paper is structured as follows. Section \ref{related-work} briefly overviews the grammar extraction mechanisms employed by other prominent open source machine translation systems. Section \ref{data-structures} reviews the data structures and algorithms that are used for efficient run-time calculation of SCFG rules. Finally, section \ref{using-joshua} provides details for configuring and running the Joshua grammar extraction tools.

@@

\section{Implementation: Data Structures and algorithms}
\label{data-structures}

\begin{figure}
\begin{algorithmic}[1]
\Function{Compare\_Elements}{$index1$,  $index2$, $max$, $corpusEnd$}
	\For {$i=0$; $i<max$; $i++$}
		\If {$index1+i < corpusEnd$ and $index2+i > corpusEnd$}
			\State \Return{$1$}
		\ElsIf {$index2+i < corpusEnd$ and $index1+i>corpusEnd$}
			\State \Return{$-1$}
		\ElsIf {$corpus[index1+i]$ is lexicographically < $corpus[index2+i]$}
			\State \Return {$-1$}
		\ElsIf {$corpus[index1+i]$ is lexicographically > $corpus[index2+i]$}
			\State \Return {$1$}
		\EndIf
	\EndFor
	\State \Return {$0$}
\EndFunction
\end{algorithmic}
\caption{}
\label{sort}
\end{figure}

@@

%\end{algorithm}
\caption{Query intersection algorithm implemented in Joshua. This algorithm is adapted from a corrected version (Lopez, p.c.) of query intersection {\protect \cite{Lopez2008}}.}
\label{query}
\end{figure}

To extract hierarchical translation rules at run-time, the decoder must have access to the aligned parallel corpus. Internally, Joshua treats all source and target words as 32-bit integers. Each unique string that is encountered is assigned a unique integer. A hash map maintains the mapping from string to integer, while a corresponding list of strings maintains the mapping from integer back to string; together, these data structures comprise the {\bf symbol table}.

A corpus can be considered a simple list whose size is equal to the number of words in the corpus. Using this approach with the symbol table, Joshua stores the source corpus as an integer array. An auxiliary array, with size equal to the number of sentences in the source corpus, is maintained; elements in this auxiliary sentence array indicate the corpus index where each sentence begins. The target corpus is likewise represented by a {\bf corpus array} and corresponding sentence array.

Once the source and target corpus arrays are available, the corresponding {\bf suffix array}s can be constructed. Given a corpus array $c$ and a symbol table, a second array is created of equal size to the corpus array. This array $s$ is initialized such that $s[x] = x]$. Where each integer in $c$ represents a word string, each integer in $s$ represents an index into $c$. The contents of $s$ are sorted, using the element comparison function defined in figure \ref{sort}. After sorting, the indices of all instances in the corpus of any given phrase are located in a contiguous segment in the suffix array $s$.

While a phrase-based decoder can simply look up any required phrase in a suffix array, hierarchical decoders must deal with discontinuous phrases that include gaps. To deal with such phrases, \namecite{Lopez2008} defines an incremental algorithm for constructing a specialized trie \cite{Fredkin1960} to represent the SCFG translation grammar. Given a source sentence, this algorithm constructs a {\bf prefix tree with suffix links} by first examining all possible contiguous source phrases, and uses the source suffix array to look up translations for contiguous phrases. Hierarchical phrases that consist of a contiguous phrase preceded or followed by a single nonterminal $X$ can be constructed directly from the corresponding contiguous phrase. In this manner, the tree is gradually constructed into a grammar containing contiguous phrases and simple hierarchical phrases.

More complex hierarchical phrases are constructed according to the {\sc Query\_Intersect} function in figure \ref{query}. This function takes two smaller phrases $a\alpha$ and $\alpha{}b$, along with the list of indices where these phrases are located. These two lists can be efficiently processed to determine the locations where the two phrases intersect to form the more complex phrase $a\alpha{}b$. In this way, all source hierarchical phrases can be located.

Each node in the prefix tree corresponds to a unique source phrase; each node stores the complete list of all indices in the source corpus where that node's phrase occurs. These locations are used in conjunction with the target corpus array and the word alignment data to construct SCFG translation rules.
